\section{Introduction}

Software defined radios (SDRs) have been around for many years. However, as the cost of SDRs
continues to drop, the technology becomes much more accessible. Additionally, open source
tools such as GNU Radio make developing for SDRs much easier. GNU Radio provides a feature rich
ecosystem that provides a wealth of signal processing blocks. Though hardware is not a direct
component in GNU Radio, numerous other developers and projects have integrated hardware functionality
into the system either natively or through additional out-of-tree modules.

GNU Radio Companion allows for GUI development of PHY and MAC layer protocols within the
GNU Radio environment.The project itself is implemented in a combination of Python and C++ modules. 
Ettus research, a division of National Instruments, created the Universal Software Radio Peripheral 
(USRP) which is the SDR we chose for the project and a popular choice among many projects.
Ettus also released the Universal Hardware Drivers (UHD) which allow for the use of the USRP with GNU Radio.   

Cognitive Radio Networks (CRNs) are networks made up of SDRs that are capable of making 
intelligent decisions on their own and adjusting parameters such as signal strength. 
Many Cognitive Radio scenarios are designed around the idea of ad-hoc or mesh
networks. In these networks, all of the associated radio components are able to talk to each
other either directly or by "hopping" from one node to another until they reach their destination. 
The Better Approach to Mobile Ad-hoc Networks (BATMAN) project created the Batman-adv protocol. 
This layer 2 protocol has a fairly large community and is integrated into the Linux Kernel. 

Our goal with this project was to create a low cost, open source platform, that could serve as a
basis for future projects in Cognitive Radio Network research. Our hope is that by combining the GNU
Radio and Batman-adv projects, more collaborative research can be done with a standarized toolset. 
Our platform will also allow for students who are looking to create application layer products that 
leverage cognitive radio meshes to get started without having to worry about all the intracasies of
these networks. 
