% Chapter 5

\chapter{Discussion} % Main chapter title

\label{Chapter5} % For referencing the chapter elsewhere, use \ref{Chapter1} 

%----------------------------------------------------------------------------------------

% Define some commands to keep the formatting separated from the content 
%\newcommand{\keyword}[1]{\textbf{#1}}
%\newcommand{\tabhead}[1]{\textbf{#1}}
%\newcommand{\code}[1]{\texttt{#1}}
%\newcommand{\file}[1]{\texttt{\bfseries#1}}
%\newcommand{\option}[1]{\texttt{\itshape#1}}

%----------------------------------------------------------------------------------------
This chapter focuses on discussing what was discovered from the creation and testing of ARCAM-Net. The limitations of the network are provided for reference. Applications that were run on the network are discussed as well. 

\section{Constraints and Limitations}



The results from our experiments show that the network is functioning as a multi hop SDRN. In addition to the experiments performed, we were also able to use Secure Shell (SSH) and Secure Copy (SCP) over multiple hops of the SDRN. However, it is clear that more work needs to be done. In a deployed network, packet loss as high as is seen in this network is not optimal. Therefore, it is important to examine ways to mitigate packet loss and increase throughput. 

For example, machine learning or artificial intelligence algorithms could be used to adjust transmission parameters as issues are detected. It is likely that a change in frequency or amplitude could mitigate some of the packet loss. For example, if the loss is due to crowding near one node, frequency hopping could be employed to shift to better operating conditions. Batctl's link quality metric, as shown in Figures \ref{fig:2Hops} and \ref{fig:NewHop} could help with finding weak nodes and making decisions. This would be the beginning of ARCAM-Net's transition from a SDRN to a Cognitive Radio Network (CRN). 

Furthermore, it would be beneficial to either improve upon A.L.F.R.E.D. or implement certain features in a new way in order to handle the frequency changing. If we have each node wait for an acknowledgment from its immediate neighbors before changing frequency, that node could then change its frequency knowing that the data will propagate to the rest of the network. 

Batctl is able to report the immediate next hop neighbors, so the program could use this information to only wait for acknowledgment from neighbors instead of waiting for the entire network to be ready to change. In order for the current A.L.F.R.E.D. setup to function, a delay was needed to give the network time to respond. Therefore, an asynchronous acknowledge would likely speed up the frequency change. 


%----------------------------------------------------------------------------------------

\section{SDR Network}

Our tests show that ARCAM-Net is a fully functioning SDR Network. We have succesfully rung ping tests, sent UDP packets, used SSH, used SCP, and even ran distributed Erlang programs over the SDRN. Thanks to the way the OSI model abstracts each layer, all programs that are designed to operate on Layer 3 or higher should see ARCAM-Net as a normal network. Though in its current state it is slower than a traditional WiFi network, there is nothing preventing any normal program from using the SDRN in the way it would use a Wi-Fi or LAN. 

%----------------------------------------------------------------------------------------

\section{Cognitive Networks}

ARCAM-Net is not a cognitive network. However, it is certainly adaptable to be used for cognitive network testing. Developing cognitive algorithms was beyond the scope of this project. Therefore, the concept was broken down into two parts: The decision making itself, and the propogation of this decision to the rest of the network. A.L.F.R.E.D. was chosen to implement the decision propogation and the web interface is meant to be used as a tool for simulating the cognitive decision making. The web interface enables the user to increase or decrease the frequency of operation. This was used to simulate the nodes themselves having cognitive decision making abilities. Using SocketIO, the node can transmit the frequency the Flask server. The Flask server will then update GNU Radio with the new frequency. This way, the logical decision making can be abstracted from the standard operation of the SDR. 

In our testing we determined A.L.F.R.E.D. is not a good fit for transmitting frequency data. But we can use whatever system eventually replaces A.L.F.R.E.D. as part of the cognitive network toolchain. When a node makes a decision to change frequency, it can use this tool to distribute the information to the rest of the network. 

\section{Applications}

All of our application layer tests showed that programs could be run over the network. SSH was able to run commands from computer to computer. SCP was used to successfully send data to and from different computers over the network. The python and erlang chat applications worked succesffully and were able to transmit messages. Erlang itself is a programming language aimed at working in distributed systems and presents a unique opportunity for future work.  
